Outbreak Threshold Detection

### 2023-02-25 notebooks/Julia/simulation-detection.qmd
Update stochastic model to use tau-leaping Gillespie Needed to change the core functions to be able to use tau-leaping as modelingtoolkit doesn't seem to handle leaping.

```{julia}
using ModelingToolkit, JumpProcesses, Random, Statistics
using DataFrames, CairoMakie

@@

```{julia}
@parameters β γ
@variables t S(t) I(t) R(t)
N = S + I + R
rate₁   = β * S * I/N
affect₁ = [S ~ S - 1, I ~ I + 1]
rate₂   = γ*I
affect₂ = [I ~ I - 1, R ~ R + 1]
j₁      = ConstantRateJump(rate₁,affect₁)
j₂      = ConstantRateJump(rate₂,affect₂)
@named js      = JumpSystem([j₁, j₂], t, [S, I, R], [β, γ])
```

```{julia}
u₀ = [S => 990, I => 10, R => 0];
p = [β => 0.5, γ => 0.25];
```

```{julia}
tmax = 250.0
tspan = (0.0, tmax)
δt = 1.0;
Random.seed!(1234);
```

```{julia}
function run_sir_sim(; js, u₀, tspan, p)
    prob = DiscreteProblem(js, u₀, tspan, p)
    sir_jprob = JumpProblem(js, prob, Direct());
    sol_jump = solve(sir_jprob, SSAStepper());
    return sol_jump(0:δt:tmax);
end
```

```{julia}
function create_sim_array(; out_jump)
    array_jump[1:3, :] = Array(out_jump)
    array_jump[4, :] .= out_jump.t

    return array_jump
end
```

@@

array_jump = zeros(Float64, 4, Int64(tmax) + 1)
all_sims_array = zeros(Float64, 4, Int64(tmax) + 1, nsims)
sim_means = zeros(Float64, 3, Int(tmax) + 1)

for i in 1:nsims
    out_jump = run_sir_sim(js = js, u₀ = u₀, tspan = tspan, p = p)
    array_jump = create_sim_array(out_jump = out_jump)

    all_sims_array[:, :, i] = array_jump

    lines!(ax, array_jump[4, :], array_jump[1, :], linewidth = 4, color = ("dodgerblue4", 0.05))
    lines!(ax, array_jump[4, :], array_jump[2, :], linewidth = 4, color = ("firebrick3", 0.05))
    lines!(ax, array_jump[4, :], array_jump[3, :], linewidth = 4, color = ("chocolate2", 0.05))

@@

for sim in 1:Int(tmax)+1, outcome in 1:3
    sim_means[outcome, sim] = mean(all_sims_array[outcome, sim, :])
end

lines!(ax, array_jump[4, :], sim_means[1, :], linewidth = 4, color = "dodgerblue4")
lines!(ax, array_jump[4, :], sim_means[2, :], linewidth = 4, color = "firebrick3")
lines!(ax, array_jump[4, :], sim_means[3, :], linewidth = 4, color = "chocolate2")

>>>>

```{julia}
using JumpProcesses, StochasticDiffEq, Random, Statistics, LinearAlgebra
using DataFrames, CairoMakie, ProgressMeter

@@

```{julia}
β = 0.1 / 1000.0
ν = 0.01;
p = (β, ν)
```

```{julia}
tmax = 250.0
δt = 0.001
tspan = (0.0, tmax)
tlength = length(tspan[1]:δt:tspan[2])
Random.seed!(1234);
```

```{julia}
function rate(out, u, p, t)
    out[1] = p[1] * u[1] * u[2]   # β * S * I
    out[2] = p[2] * u[2]          # ν * I
    nothing
end
```

```{julia}
c = zeros(3, 2)
# S + I --> I
c[1, 1] = -1    # S -> S - 1
c[2, 1] = 1     # I -> I + 1

# I --> R
c[2, 2] = -1    # I -> I - 1
c[3, 2] = 1     # R -> R + 1
```

```{julia}
function change(du, u, p, t, counts, mark)
    mul!(du, c, counts)
    nothing
end
```

```{julia}
rj = RegularJump(rate, change, 2)

u₀ = [1000.0, 50.0, 0.0]
prob = DiscreteProblem(u₀, tspan, p)
jump_prob = JumpProblem(prob, Direct(), rj)

sol = solve(jump_prob, TauLeaping(); dt = δt)
```

```{julia}
sir_array = zeros(Float64, 4, tlength) 
sir_array[1:3, :] = Array(sol)
sir_array[4, :] = sol.t
```

```{julia}
f = Figure()
ax = Axis(
    f[1, 1],
    xlabel = "Time",
    ylabel = "Number"
    )

lines!(ax, sir_array[4, :], sir_array[1, :])
lines!(ax, sir_array[4, :], sir_array[2, :])
lines!(ax, sir_array[4, :], sir_array[3, :])

f
```

```{julia}
function run_sir_sim(; u₀, tspan, p)
    prob = DiscreteProblem(u₀, tspan, p)
    jump_prob = JumpProblem(prob, Direct(), rj)
    sol_jump = solve(jump_prob, TauLeaping(); dt = δt)
    return sol_jump(0:δt:tmax);
end
```

```{julia}
function create_sim_array(; out_jump)
    test_array[1:3, :] = Array(out_jump)
    test_array[4, :] .= out_jump.t

    return test_array
end
```

@@

sir_array = zeros(Float64, 4, tlength)
all_sims_array = zeros(Float64, 4, tlength, nsims)
sim_means = zeros(Float64, 3, tlength)

@showprogress for i in 1:nsims
    out_jump = run_sir_sim(u₀ = u₀, tspan = tspan, p = p)
    sir_array = create_sim_array(out_jump = out_jump)

    all_sims_array[:, :, i] = sir_array

    lines!(ax, sir_array[4, :], sir_array[1, :], linewidth = 4, color = ("dodgerblue4", 0.01))
    lines!(ax, sir_array[4, :], sir_array[2, :], linewidth = 4, color = ("firebrick3", 0.01))
    lines!(ax, sir_array[4, :], sir_array[3, :], linewidth = 4, color = ("chocolate2", 0.01))

@@

@showprogress for sim in 1:tlength, outcome in 1:3
    sim_means[outcome, sim] = mean(all_sims_array[outcome, sim, :])
end

lines!(ax, sir_array[4, :], sim_means[1, :], linewidth = 3, color = "dodgerblue4")
lines!(ax, sir_array[4, :], sim_means[2, :], linewidth = 3, color = "firebrick3")
lines!(ax, sir_array[4, :], sim_means[3, :], linewidth = 3, color = "chocolate2")
